---
title: "ST406 - Project 2 Data, EDA"
author: "P.I.N.Kehelbedda"
date: "1/3/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    number_sections: true
    df_print: kable
---

```{r Libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(readr) # read data
library(ggplot2) # plot data
library(dplyr) # data manipulation
# library(tidyverse) # data manipulation 2.0
library(stringr) # str manipulation
library(GGally) # ggpairs
library(openair) # plot wind data
library(lubridate) # date, extract months, days
library(gridExtra)
library(Amelia)
library(corrplot)
library(ggfortify)
library(RColorBrewer)
library(forecast)
```

# Introduction

# Data

This data set is meteorological data from the HI-SEAS weather station from four months (September through December 2016) between Mission IV and Mission V.

The data set contains such columns as: "wind direction", "wind speed", "humidity" and temperature. The response parameter that is to be predicted is: "Solar_radiation". It contains measurements for the past 4 months and you have to predict the level of solar radiation.
Just imagine that you've got solar energy batteries and you want to know will it be reasonable to use them in future?.

## Reading data

```{r, message=FALSE}
SolarRadPrediction <- read_csv("Data/SolarRadPrediction.csv", 
    col_types = cols(Data = col_datetime(format = "%m/%d/%Y %H:%M:%S %p"), 
        Time = col_time(format = "%H:%M:%S"), 
        TimeSunRise = col_time(format = "%H:%M:%S"), 
        TimeSunSet = col_time(format = "%H:%M:%S")))

colnames(SolarRadPrediction)[8] <- "WindDirection"
dim(SolarRadPrediction)
str(SolarRadPrediction)
head(SolarRadPrediction)
```

## About data

Data set contains these variables,

+ UNIXTIME: Unix form of time variable.
+ Data: Date in format of yyyy-mm-dd
+ Time: The local time in the format of hh:mm:ss 24-hr.
+ Radiation: Solar radiation in watts per meter squared ($1kg/s^3$).
+ Temperature: Temperature in degrees fahrenheit ($^\circ F$).
+ Pressure: Barometric Pressure in $Hg$.
+ Humidity: Humidity precent.
+ WindDirection: Wind derection in degrees.
+ Speed: Wind speed in miles per hour (mph).
+ TimeSunRise: Hawaii time of Sun rise.
+ TimeSunSet: Hawaii time of Sun set.

I'm assuming that the location is Hawaii. Furthermore, the wind direction is measured clockwise from 0 degrees North.

## Data Pre-Processing

### Missing values



```{r}
mean(is.na(SolarRadPrediction))
```

There is no missing values. Data is 100% cleaned.


### Dates


# EDA

## Data

```{r}
df <- SolarRadPrediction[, -1]
colnames(df)[1] <- "Date"
colnames(df)[8] <- "WindSpeed"
```


## Dependent variable Radiation

### Radiation

#### Tables

```{r}
summary(df$Radiation)

## Radiation highest value when time ?
df %>%
  filter(Radiation == max(Radiation))
## Min radiation
df %>%
  filter(Radiation == min(Radiation))
```

#### Figures

```{r}
ggplot(data = df, aes(x = Time, y = Radiation)) +
  geom_line(col = "blue") + labs(title = "Radiation by Time")
```

## Handle Date variable

```{r}
# GetMonths 
m <- month(df$Date)
table(m)
# Get dates
d <- day(df$Date)
```


## Additional Tables

```{r}
summary(df[, c(3,4,5,6,8)])
```


## Additional Figures

```{r}
# Plotting wind speed and wind direction over time
df1 <- df %>%
  select(ws = WindSpeed, wd = WindDirection, date = Date)
## weekdays wind
polarFreq(mydata = df1, cols = "jet")
polarFreq(mydata = df1, cols = "jet", type = "weekday")
## weekdays * season wind
# polarFreq(mydata = df1, cols = "jet", type = c("weekday", "season"))

#Correlations
#numData <- df[, -c(1,2,9,10)]
#ggpairs(numData)
```


```{r}
dfd <- df # new data frame since no need to change df

month <- month(df$Date)
day <- day(df$Date)

ggplot(dfd, aes(factor(month), Radiation)) + 
  geom_boxplot(aes(fill = factor(month))) +
  ggtitle("Boxplot of Radiation values for each month") +
  scale_x_discrete(labels = c("September", "Octomber", "November", "December")) +
  scale_fill_discrete(name = "Months", 
                      labels = c("September", "Octomber", "November", "December")) +
  xlab("Month")
```

```{r functions}
# df <- SolarRadPrediction
# # any(is.na(df))
# 
# getDate<-function(x,pos1,pos2){
#   if(pos1==1){
#     val<-as.numeric(strsplit(strsplit(as.character(x)," ")[[1]][pos1],'/')[[1]][pos2])
#   }
#   else if(pos1==3 & pos2==0){
#     val<-as.factor(strsplit(strsplit(as.character(x)," ")[[1]][pos1],'/')[[1]])
#   }
#   return(val)
# }
# 
# getTIME<-function(x,pos){
#   val<-strsplit(as.character(x),":")[[1]][pos]
#   return(as.numeric(val))
# }
```


```{r}
# df$Month <- sapply(df$Data, getDate,1,1)
# df$Day <- sapply(df$Data,getDate,1,2)
# df$Year <- sapply(df$Data,getDate,1,3)
# df$TimeAbbr <- sapply(df$Data,getDate,3,0)
# df$hour <- sapply(df$Time,getTIME,1)
# df$minute <- sapply(df$Time,getTIME,2)
# df$sec <- sapply(df$Time,getTIME,3)
```



```{r}
# mymonths <- c("January","February",
#               "March","April","May","June","July",
#               "August","September","October","November","December")
# 
# df$MonthAbb <- mymonths[ df$Month ]
# df$ordered_month <- factor(df$MonthAbb, levels = month.name)
# 
# df$DateTs<-as.POSIXct(paste0(df$Year,'-',
#                              df$Month,'-',
#                              df$Day,' ',
#                              as.character(df$Time)),
#                       format="%Y-%m-%d %H:%M:%S")
# 
# df$DailyTs <- as.POSIXct(as.character(df$Time), format="%H:%M:%S")
# 
# df$DiffTime<-as.numeric(difftime(as.POSIXct(paste0(df$Year,
#                                                    '-',df$Month,
#                                                    '-',df$Day,' ',
#                                                    as.character(df$TimeSunSet)), 
#                                             format="%Y-%m-%d %H:%M:%S"),
#   
#                                  as.POSIXct(paste0(df$Year,'-',
#                                                    df$Month,'-',
#                                                    df$Day,' ',
#                                                    as.character(df$TimeSunRise)), 
#                                             format="%Y-%m-%d %H:%M:%S"),
#                                  units='sec'))
```



```{r}
# plot
# ggplot(data=df,aes(x=Radiation,fill=ordered_month)) + 
#   geom_histogram(bins=100) + 
#   scale_y_log10() + 
#   scale_fill_manual(name="",values=rainbow(4)) + 
#   theme(legend.position='top') + 
#   facet_wrap(~ordered_month) +
#   xlab('Radiation level [W/m^-2]') + ylab('Count')
# 
# # plot
# ggplot(data=df,aes(x=DiffTime,y=Radiation)) + 
#   geom_point(aes(color=ordered_month)) +
#   scale_color_manual(name="",values=rainbow(4)) + 
#   theme(legend.position='top') + xlab("SunSet -SunRise [sec]")
# 
# # plot
# df %>% select(ordered_month,Day,Radiation) %>% 
#   group_by(ordered_month,Day) %>% 
#   summarise(dailyRad = mean(Radiation)) %>% 
#   ggplot(aes(x=ordered_month,y=dailyRad,color=dailyRad)) + 
#   scale_color_gradientn(colours=rev(brewer.pal(10,'Spectral'))) + 
#   geom_boxplot(colour='black',size=.4,alpha=.5) + 
#   geom_jitter(shape=16,width=.2,size=2) +
#   xlab('') + ylab('') + theme(legend.position='top')
```

